30 years of privacy research for the masses \& \newline 15 years of decentralization attempts

### 2014-03-14 paper.tex
Added the missing cites

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	There are quite a few distributed anonymous communication networks such as Tor, yet when it comes to decentralized anonymous systems the numbers are rather small and certain problems arise. In this technical report we explain how Tor works and which problems it currently faces. After that, we talk about why would like to decentralize peer-to-peer networks, what the difficulties are with decentralizing networks and show which projects have made an attempt to decentralize or are already (partially) decentralized. Finally, we provide a table in where we compare these projects against each other, which will also be public available on the Github page mention in this report. (We are going to mention it, right?)
\end{abstract}

@@

		Before a circuit can be established, a path has to be chosen. The current path selection algorithm in Tor selects nodes based on the bandwidth of the nodes [link naar cacr2011-20]. Nodes that have more bandwidth, have a higher probability to be chosen for the circuit setup, however, the same node can't be used more than once in a circuit.

@@

		While Tor guarantees anonymity, there are some disadvantages using it. The main disadvantage is that the Tor network is slow. According to the Tor Metrics project [link naar tor metrics project], it takes about 6 seconds to download 1 MiB of data.

@@

		First of all, the congestion control does not work well. The network has some problems handling bulk transfers, such as when downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol for links between Tor relays. Goldberg et al. [linkje] have proposed PCTCP which could improve the response time of the Tor network with 60\% and the download time of files with 30\%.

@@

		As we saw, the directory service generates overhead on the network. If we could replace this central authority, we could reduce the overhead of the network and improve the performance. This means, we would like to have Tor decentralized. Many research has been done already on the decentralization of Tor [linkjezz] but so far, no anonymous, well-performing decentralized network has come up. This is due to the fact that there are many problems decentralization brings along with.

@@

	\subsection{Low-Resource Routing attacks}
		Another possible attack on the anonimity of tor is the so called \emph{low-resource routing attack} [linkje naar paper]. With this attack, it is possible for an adversary to perform an end-to-end traffic analysis with minimal resourcing, thus comprimising the anonimy Tor provides. The idea of this attack is that a malicious Onion router can lie about it's bandwidth, thus advertising a much higher bandwidth than it actually has. Because of the Tor path selection algorithm that prefers high-bandwidth nodes, the chance that a malicious entry node and exit node are chosen, is high.

@@

		A fully distributed network that implements such a solution is the Bitcoin network [BITCOIN PAPER: http://s.kwma.kr/pdf/Bitcoin/bitcoin.pdf], in which computing power, and not the number of nodes is important for the general network consensus.

	\subsection{Replay Attacks}
		A replay attack [cite naar replay attack paper] happens when a malicious entry node duplicates cells and send them again. Since Tor uses uses the counter mode of Advanced Encryption Standard (AES-CTR) for encryption and decryption, the counter will be wrong when the duplicated package arrives causing the connection to be torn down.

@@

	As we saw in section \ref{sec:tor}, we would like to have a decentralized anonymous network. This isn't an easy task: 15 years of decentralization attempts on the Tor network are behind us. Several well-used decentralized peer-to-peer overlay networks are active such as Gnutella [link], Freenet [link] and Tapestry. These networks will be described in more detail in the next chapter. Among these networks are some overlay anonymous networks such as Tarzan.

@@

		There are several techniques to add an entry to the routers NAT table. Universal Plug and Play (UPnP) is one of those techniques, where the local device uses a HTTP request to the router to associate a port with the devices private network address. This technique is not available everywhere and sometimes considered a security risk. Different implementations of NAT require different techniques, such as \textit{hole punching}, \textit{relaying} or \textit{reversal}, as described by Wocker et al. [http://www.pap.vs.uni-due.de/files/wacker-nat-traversal.pdf].

@@

		Tor currently uses an interactive forward-secret key-exchange protocol called the \emph{Tor Authentication Protocol} (TAP) [linkje naar ace]. This protocol uses telescoping which means that the initiator negotiates session keys with each successive hop in the circuit. There are several proposals for more efficient key exchange methods. One of them is \emph{ACE}, an one-way authenticated key exchange protocol. The authors of this methods claim to have an 46\% efficiency improvement on the side of the client and an efficiency improvement of nearly 19\% on the side of the Onion Routers. ACE requires clients to send one extra element in the key exchange. This however, does not introduce any overhead because the element fits in the unused space in a cell.

@@

	Many other decentralized peer-to-peer networks have been proposed and some have been implemented [linkjes naar papers]. Examples of fully decentralized networks are Gnutella, Freenet, Pastry, Tapestry, and Tarzan. Each of these decentralized networks will be discussed in the next section. Note that we're only considering fully decentralized networks here. Networks that are partially or hybrid centralized, such as KaZaA or Napster, are not discussed here.[link naar survey p2p?]

@@

	\subsection{Tapestry} % 2001
		Tapestry is based on the Plaxton mesh data structure [linkje naar plaxton paper], which maintains pointers to node in the network whose IDs match the elements of a tree-like structure of ID prefixes up to a digit position. A property of Tapestry is that it offers load distribution and routing locality.

@@

		Tapestry is a fundamental component of OceanStore [linkje naar paper over OceanStore], a decentralized storage system. Tapestry is also used in systems such as Bayeux and SpamWatch, a decentralized spam-filtering system.

@@

	\subsection{Tarzan} % 2002
		Tarzan [TARZAN PAPER] is a fully distributed peer-to-peer anonymity network. It implements a network address translator (NAT) to bridge between nodes running Tarzan and the internet. This means that services don\'t have to be aware of the fact they are running through Tarzan.

		Tarzan requires knowledge of a few existing nodes to bootstrap and uses a gossiping protocol to discover other nodes. Nambiar et al. showed however that this does not scale beyond roughly 10,000 nodes [NISAN, SALSA PAPER].

@@

	\subsection{Torsk} % 2009
		Torsk [TORSKKKK PAPER] is an extension to Tor, designed to be an interoperable replacement for the circuit creation and Directory Service as used by Tor. The authors argue that the current Directory Service does not scale, with the percentage of the traffic in a network dedicated to node discovery growing as the number of nodes grow. With the 2009 version of Tor, they argue that 100\% of the networks traffic would consist of node discovery traffic with roughly 1.2 million clients.

@@

	\subsection{Other decentralized systems}
		Some other decentralized peer-to-peer systems has been developed, focussing on anonymity. An example of such a system is AP3 (Anonymizing Peer-to-Peer Proxy) [linkje naar AP3 paper] which makes cooperative, decentralized anonymous communication possible. The AP3 system provides clients with three primitives: anonymous message delivery, anonymous channels and secure pseudonyms. Users are able to send and receive unicast, multicast and anycast messages anonymously. The strategy that AP3 is using for message delivery is similar to that of Tarzan: it relies on a network of peers to forward messages. A node along the request path, does not know whether the node from which it receives a message is the message's originator or simply another forwarding peer.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	There are quite a few distributed anonymous communication networks such as Tor, yet when it comes to decentralized anonymous systems the numbers are rather small and certain problems arise. In this technical report we explain how Tor works and which attacks are possible and which vulnerabilities currently exist in Tor. After that, we talk about why would like to decentralize peer-to-peer networks, what the difficulties are with decentralizing networks and show which projects have made an attempt to decentralize or are already (partially) decentralized. Finally, we provide a table in where we compare these projects against each other, which will also be public available on the Github page mention in this report.
\end{abstract}

@@

		Before a circuit can be established, a path has to be chosen. The current path selection algorithm in Tor selects nodes based on the bandwidth of the nodes \cite{wang2012congestion}. Nodes that have more bandwidth, have a higher probability to be chosen for the circuit setup, however, the same node can't be used more than once in a circuit.

@@

		While Tor guarantees anonymity, there are some disadvantages using it. The main disadvantage is that the Tor network is slow. According to the Tor Metrics project \cite{tormetricsprojectwebsite}, it takes about 6 seconds to download 1 MiB of data.

@@

		First of all, the congestion control does not work well. The network has some problems handling bulk transfers, such as when downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol for links between Tor relays. Goldberg et al. \cite{alsabah2013pctcp} have proposed PCTCP which could improve the response time of the Tor network with 60\% and the download time of files with 30\%.

@@

		As we saw, the directory service generates overhead on the network. If we could replace this central authority, we could reduce the overhead of the network and improve the performance. This means, we would like to have Tor decentralized. Many research has been done already on the decentralization of Tor \cite{wang2012octopus, alsabah2013pctcp} (WHICH CITES CAN BE ADDED HERE? A LOT ARE MORE SUITEABLE FOR THE CURRENT DECENTRALIZED SYSTEM SECTION) but so far, no anonymous, well-performing decentralized network has come up. This is due to the fact that there are many problems decentralization brings along with.

@@

	\subsection{Low-Resource Routing attacks}
		Another possible attack on the anonimity of tor is the so called \emph{low-resource routing attack} \cite{bauer2007low}. With this attack, it is possible for an adversary to perform an end-to-end traffic analysis with minimal resourcing, thus comprimising the anonimy Tor provides. The idea of this attack is that a malicious Onion router can lie about it's bandwidth, thus advertising a much higher bandwidth than it actually has. Because of the Tor path selection algorithm that prefers high-bandwidth nodes, the chance that a malicious entry node and exit node are chosen, is high.

@@

		A fully distributed network that implements such a solution is the Bitcoin network \cite{nakamoto2008bitcoin}, in which computing power, and not the number of nodes is important for the general network consensus.

	\subsection{Replay Attacks}
		A replay attack \cite{pries2008new} happens when a malicious entry node duplicates cells and send them again. Since Tor uses uses the counter mode of Advanced Encryption Standard (AES-CTR) for encryption and decryption, the counter will be wrong when the duplicated package arrives causing the connection to be torn down.

@@

	As we saw in section \ref{sec:tor}, we would like to have a decentralized anonymous network. This isn't an easy task: 15 years of decentralization attempts on the Tor network are behind us. Several well-used decentralized peer-to-peer overlay networks are active such as Gnutella \cite{ripeanu2001peer}, Freenet \cite{lua2005survey} and Tapestry. These networks will be described in more detail in the next chapter. Among these networks are some overlay anonymous networks such as Tarzan.

@@

		There are several techniques to add an entry to the routers NAT table. Universal Plug and Play (UPnP) is one of those techniques, where the local device uses a HTTP request to the router to associate a port with the devices private network address. This technique is not available everywhere and sometimes considered a security risk. Different implementations of NAT require different techniques, such as \textit{hole punching}, \textit{relaying} or \textit{reversal}, as described by Wocker et al. \cite{wacker2008nat}.

@@

		Tor currently uses an interactive forward-secret key-exchange protocol called the \emph{Tor Authentication Protocol} (TAP) \cite{backes2012ace}. This protocol uses telescoping which means that the initiator negotiates session keys with each successive hop in the circuit. There are several proposals for more efficient key exchange methods. One of them is \emph{ACE}, an one-way authenticated key exchange protocol. The authors of this methods claim to have an 46\% efficiency improvement on the side of the client and an efficiency improvement of nearly 19\% on the side of the Onion Routers. ACE requires clients to send one extra element in the key exchange. This however, does not introduce any overhead because the element fits in the unused space in a cell.

@@

	Many other decentralized peer-to-peer networks have been proposed and some have been implemented \cite{mislove2004ap3, rennhard2002introducing, panchenko2006nisan, rowstron2001pastry, nambiar2006salsa, freedman2002tarzan, ripeanu2001peer, androutsellis2004survey}. Examples of fully decentralized networks are Gnutella, Freenet, Pastry, Tapestry, and Tarzan. Each of these decentralized networks will be discussed in the next section. Note that we're only considering fully decentralized networks here. Networks that are partially or hybrid centralized, such as KaZaA or Napster, are not discussed here.

@@

	\subsection{Tapestry} % 2001
		Tapestry is based on the Plaxton mesh data structure, which maintains pointers to node in the network whose IDs match the elements of a tree-like structure of ID prefixes up to a digit position. A property of Tapestry is that it offers load distribution and routing locality.

@@

		Tapestry is a fundamental component of OceanStore, a decentralized storage system. Tapestry is also used in systems such as Bayeux and SpamWatch, a decentralized spam-filtering system.

@@

	\subsection{Tarzan} % 2002
		Tarzan \cite{freedman2002tarzan} is a fully distributed peer-to-peer anonymity network. It implements a network address translator (NAT) to bridge between nodes running Tarzan and the internet. This means that services don\'t have to be aware of the fact they are running through Tarzan.

		Tarzan requires knowledge of a few existing nodes to bootstrap and uses a gossiping protocol to discover other nodes. Nambiar et al. showed however that this does not scale beyond roughly 10,000 nodes \cite{panchenko2009nisan, nambiar2006salsa}.

@@

	\subsection{Torsk} % 2009
		Torsk \cite{mclachlan2009scalable} is an extension to Tor, designed to be an interoperable replacement for the circuit creation and Directory Service as used by Tor. The authors argue that the current Directory Service does not scale, with the percentage of the traffic in a network dedicated to node discovery growing as the number of nodes grow. With the 2009 version of Tor, they argue that 100\% of the networks traffic would consist of node discovery traffic with roughly 1.2 million clients.

@@

	\subsection{Other decentralized systems}
		Some other decentralized peer-to-peer systems has been developed, focussing on anonymity. An example of such a system is AP3 (Anonymizing Peer-to-Peer Proxy) \cite{mislove2004ap3} which makes cooperative, decentralized anonymous communication possible. The AP3 system provides clients with three primitives: anonymous message delivery, anonymous channels and secure pseudonyms. Users are able to send and receive unicast, multicast and anycast messages anonymously. The strategy that AP3 is using for message delivery is similar to that of Tarzan: it relies on a network of peers to forward messages. A node along the request path, does not know whether the node from which it receives a message is the message's originator or simply another forwarding peer.
